---
title: "Report Project MRR 2022 - Ocean - Binomial 18"
output:
  pdf_document: default
  html_document: default
date: "2022-11-14"
---

Do Thanh Dat LE

Gan WANG

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
```

```{r}
library(readxl)
library(reshape2)
library(ggplot2)
```

# Read data

```{r}
Vect = read_excel("Vect.xlsx")
depth = read_excel("depth.xlsx")
names(Vect) = gsub("'","", names(Vect))
```

# Draw histogram of all CHL

```{r}
# histogram before tranform
ggplot(melt(Vect[,c(23:40)]),aes(x = value),) + 
  facet_wrap(~variable,scales = "free_x") + 
  geom_histogram() + ggtitle("Histogram of all CHL") 
# histogram after transform to log10CHL
ggplot(melt(log10(Vect[,c(23:40)])),aes(x = value)) + 
  facet_wrap(~variable,scales = "free_x") + 
  geom_histogram()
```

# Transform the CHL to log10CHL

```{r}
# Transform the CHL to log10CHL
Vect$log10CHL2 = log10(Vect$`CHL 2`)
Vect$log10CHL3 = log10(Vect$`CHL 3`)
Vect$log10CHL4 = log10(Vect$`CHL 4`)
Vect$log10CHL5 = log10(Vect$`CHL 5`)
Vect$log10CHL6 = log10(Vect$`CHL 6`)
Vect$log10CHL7 = log10(Vect$`CHL 7`)
Vect$log10CHL8 = log10(Vect$`CHL 8`)
Vect$log10CHL9 = log10(Vect$`CHL 9`)
Vect$log10CHL10 = log10(Vect$`CHL 10`)
Vect$log10CHL11 = log10(Vect$`CHL 11`)
Vect$log10CHL12 = log10(Vect$`CHL 12`)
Vect$log10CHL13 = log10(Vect$`CHL 13`)
Vect$log10CHL14 = log10(Vect$`CHL 14`)
Vect$log10CHL15 = log10(Vect$`CHL 15`)
Vect$log10CHL16 = log10(Vect$`CHL 16`)
Vect$log10CHL17 = log10(Vect$`CHL 17`)
Vect$log10CHL18 = log10(Vect$`CHL 18`)
```

# Create the BATS data

```{r}
bats = Vect[(Vect$longitude>=-65)&(Vect$longitude<=-63)&(Vect$latitude>=31)&(Vect$latitude<=33),]
bats_data = bats[,-c(6:22, 24:40, 43:44)]
bats_data$sin.5days = sin(2*pi*(bats_data$`5days`/73))
bats_data$cos.5days= cos(2*pi*(bats_data$`5days`/73))
bats_data = bats_data[,-c(7)]
sin.5days = bats_data$sin.5days
cos.5days = bats_data$cos.5days
year = bats_data$year
log10CHL = bats_data[,c(8:24)] #extract 17 CHLs for 17 depths
```

# Create the spatial data

```{r}
df = Vect[,-c(6:22, 24:42, 45:61)]
for (i in 1:9) {
  location = df[(df$latitude == df$latitude[i])&(df$longitude == df$longitude[i]),]
  if ((location$latitude[1] == bats$latitude[1]) & (location$longitude[1] == bats$longitude[1])) {
    names(location) = paste(names(location),"_BATS")
  } else {
    names(location) = paste(names(location),"_",i)
  }
  location = location[, -c(7:8)]
  if (i==1) {
    spatial_data = cbind(location)
  } else {
    spatial_data = cbind(spatial_data, location) 
  }
}

spatial_data = cbind(spatial_data, sin.5days, cos.5days, year, log10CHL)
```

# Create the temporal data

```{r}
# create NA data for new variable
before.5days = data.frame(matrix(data = NA, nrow = 1241, ncol = ncol(bats_data[,-c(7:26)])))
before.10days = data.frame(matrix(data = NA, nrow = 1241, ncol = ncol(bats_data[,-c(7:26)])))
after.5days = data.frame(matrix(data = NA, nrow = 1241, ncol = ncol(bats_data[,-c(7:26)])))
after.10days = data.frame(matrix(data = NA, nrow = 1241, ncol = ncol(bats_data[,-c(7:26)])))
# name the variables
names(before.5days) = paste(names(bats_data[,-c(7:26)]),".before.5days")
names(before.10days) = paste(names(bats_data[,-c(7:26)]),".before.10days")
names(after.5days) = paste(names(bats_data[,-c(7:26)]),".after.5days")
names(after.10days) = paste(names(bats_data[,-c(7:26)]),".after.10days")
# create temporal data
for (i in 18:(nrow(bats_data))) {
  before.5days[i,] = bats_data[i-17,-c(7:26)]
  before.10days[i,] = before.5days[i-17,]
  after.5days[i-17,] = bats_data[i,-c(7:26)]
  after.10days[i-17,] = bats_data[i+17,-c(7:26)]
}
temporal_data = cbind(bats_data, before.5days, before.10days, after.5days, after.10days)
temporal_data = na.omit(temporal_data)
```

#Correlation

```{r}
cordata1 = cor(bats_data)
cordata2 = cor(spatial_data)
cordata3 = cor(temporal_data)
corrplot::corrplot(cordata1, method ="circle", tl.cex = 0.5)
corrplot::corrplot(cordata2, method ="circle", tl.cex = 0.5)
corrplot::corrplot(cordata3, method ="circle", tl.cex = 0.5)
```

# Split the data into training set, test set and validation set

## BATS data

```{r}
# sort data by year
bats_data = bats_data[order(bats_data$year),]
train_ind = as.integer((nrow(bats_data)*11)/17)
bats_train = bats_data[1:train_ind,]
pseudo = bats_data[(train_ind + 1):nrow(bats_data),]
test_id = as.integer(nrow(pseudo)/2)
bats_test = pseudo[1:test_id,]
bats_valid = pseudo[(test_id+1):nrow(pseudo),]
```

## Spatial data

```{r}
spatial_data = spatial_data[order(spatial_data$year),]
train_ind = as.integer((nrow(spatial_data)*11)/17)
spatial_train = spatial_data[1:train_ind,]
pseudo = spatial_data[(train_ind + 1):nrow(spatial_data),]
test_id = as.integer(nrow(pseudo)/2)
spatial_test = pseudo[1:test_id,]
spatial_valid = pseudo[(test_id+1):nrow(pseudo),]
```

## Temporal Data

```{r}
temporal_data = temporal_data[order(temporal_data$year),]
train_ind = as.integer((nrow(temporal_data)*11)/17)
temporal_train = temporal_data[1:train_ind,]
pseudo = temporal_data[(train_ind + 1):nrow(temporal_data),]
test_id = as.integer(nrow(pseudo)/2)
temporal_test = pseudo[1:test_id,]
temporal_valid = pseudo[(test_id+1):nrow(pseudo),]
```

# Build model to predict CHL4

## Model 1: Linear regression model with stepwise on BATS data

```{r}
#BASELINE MODEL
reg.bats = lm(log10CHL4~., data = bats_train[,-c(7:9,11:24)])
summary(reg.bats)
#STEPWISE
regstep.bats = step(reg.bats, direction = "both", trace = 0)
summary(regstep.bats)
#Predict on the test set and compute RMSE
bats_pred = predict(regstep.bats, newdata = bats_test[,-c(7:24)])
rmse_bats.reg = sqrt(mean((10^(bats_pred)-10^(bats_test$log10CHL4))^2))
rmse_bats.reg
plot(10^(bats_pred), 10^(bats_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 2: Lasso regression model (L1) on BATS data

```{r}
#LASSO MODEL
library(glmnet)
X = as.matrix(bats_train[,-c(7:24)])
Y = as.matrix(bats_train$log10CHL4)
lasso.mod = cv.glmnet(X, Y, alpha = 1, lambda = seq(0, 10, 0.01), nfolds = 10)
plot(lasso.mod)
plot(lasso.mod$glmnet.fit, "lambda", label = T)
lambda.lasso = lasso.mod$lambda.min
lasso.bats = glmnet(X, Y, alpha = 1, lambda = lambda.lasso)
coef(lasso.bats)
# Predict and compute MRSE
bats_pred_train = predict(lasso.bats, as.matrix(bats_train[,-c(7:24)]))
bats_pred.lasso = predict(lasso.bats, as.matrix(bats_test[,-c(7:24)]))
rmse_bats.lasso = sqrt(mean((10^(bats_pred.lasso)-10^(bats_test$log10CHL4))^2))
rmse_bats.lasso
R_squared.lasso = (sd(bats_pred_train))^2/(sd(bats_train$log10CHL4))^2
R_squared.lasso
adj_R.lasso = 1-(1-R_squared.lasso)*((nrow(bats_train)-1)/(nrow(bats_train)-8))
adj_R.lasso
plot(10^(bats_pred.lasso), 10^(bats_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 3: Ridge regression model (L2) on BATS data

```{r}
#Ridge model 
library(MASS)
ridge.mod = lm.ridge(log10CHL4~., bats_train[,-c(7:9,11:24)], lambda = seq(0,10,0.01))
plot(ridge.mod)
plot(ridge.mod$GCV)
best.lambda = as.numeric(names(which.min(ridge.mod$GCV)))
ridge.bats = lm.ridge(log10CHL4~., bats_train[,-c(7:9,11:24)], lambda = best.lambda)
coef(ridge.bats)
#Predict and compute RMSE, R^2 and adjusted R^2
bats_pred_train = as.matrix(cbind(rep(1,nrow(bats_train)),bats_train[,-c(7:24)]))%*%as.vector(coef(ridge.bats))
bats_pred.ridge = as.matrix(cbind(rep(1,nrow(bats_test)),bats_test[,-c(7:24)]))%*%as.vector(coef(ridge.bats))
rmse_bats.ridge = sqrt(mean((10^(bats_pred.ridge)-10^(bats_test$log10CHL4))^2))
rmse_bats.ridge
R_squared.ridge = (sd(bats_pred_train))^2/(sd(bats_train$log10CHL4))^2
R_squared.ridge
adj_R.ridge = 1-(1-R_squared.ridge)*((nrow(bats_train)-1)/(nrow(bats_train)-8))
adj_R.ridge
plot(10^(bats_pred.ridge), 10^(bats_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 4: Linear regression with spatial data

```{r}
#BASELINE MODEL
reg.spatial = lm(log10CHL4~., data = spatial_train[,-c(57:59, 61:74)])
summary(reg.spatial)
#STEPWISE
regstep.spatial = step(reg.spatial, direction = "both", trace = 0)
summary(regstep.spatial)
#Predict and compute RMSE
spatial_pred = predict(regstep.spatial, newdata = spatial_test[,-c(57:74)])
rmse_spatial.reg = sqrt(mean((10^(spatial_pred)-10^(spatial_test$log10CHL4))^2))
rmse_spatial.reg
plot(10^(spatial_pred), 10^(spatial_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 5: Linear regression with temporal data

```{r}
#BASELINE MODEL
reg.temporal = lm(log10CHL4~., data = temporal_train[,-c(7:9, 11:24)])
summary(reg.temporal)
#STEPWISE
regstep.temporal = step(reg.temporal, direction = "both", trace = 0)
summary(regstep.temporal)
#Predict and compute RMSE
temporal_pred = predict(regstep.temporal, newdata = temporal_test[,-c(7:24)])
rmse_temporal.reg = sqrt(mean((10^(temporal_pred)-10^(temporal_test$log10CHL4))^2))
rmse_temporal.reg
plot(10^(temporal_pred), 10^(temporal_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 6: Lasso regression with spatial data

```{r}
#LASSO MODEL
library(glmnet)
X = as.matrix(spatial_train[,-c(57:74)])
Y = as.matrix(spatial_train$log10CHL4)
lasso.mod = cv.glmnet(X, Y, alpha = 1, lambda = seq(0, 10, 0.01), nfolds = 10)
plot(lasso.mod)
plot(lasso.mod$glmnet.fit, "lambda", label = T)
lambda.lasso = lasso.mod$lambda.min
lasso.spatial = glmnet(X, Y, alpha = 1, lambda = lambda.lasso)
coef(lasso.spatial)
# Predict and compute MRSE
spatial_pred = predict(lasso.spatial, as.matrix(spatial_train[,-c(57:74)]))
spatial_pred.lasso = predict(lasso.spatial, as.matrix(spatial_test[,-c(57:74)]))
rmse_spatial.lasso = sqrt(mean((10^(spatial_pred.lasso)-10^(spatial_test$log10CHL4))^2))
rmse_spatial.lasso
R_squared.lasso = (sd(spatial_pred))^2/(sd(spatial_train$log10CHL4))^2
R_squared.lasso
adj_R.lasso = 1-(1-R_squared.lasso)*((nrow(spatial_train)-1)/(nrow(spatial_train)-56))
adj_R.lasso
plot(10^(spatial_pred.lasso), 10^(spatial_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 7: Ridge regression with spatial data

```{r}
#Ridge model 
library(MASS)
ridge.mod = lm.ridge(log10CHL4~., spatial_train[, -c(57:59, 61:74)], lambda = seq(0,10,0.01))
plot(ridge.mod)
best.lambda = as.numeric(names(which.min(ridge.mod$GCV)))
ridge.spatial = lm.ridge(log10CHL4~., spatial_train[, -c(57:59, 61:74)], lambda = best.lambda)
coef(ridge.spatial)
#Predict and compute RMSE
spar_pred = as.matrix(cbind(rep(1,nrow(spatial_train)),spatial_train[,-c(57:74)]))%*%as.vector(coef(ridge.spatial))
spatial_pred.ridge = as.matrix(cbind(rep(1,nrow(spatial_test)),spatial_test[,-c(57:74)]))%*%as.vector(coef(ridge.spatial))
rmse_spatial.ridge = sqrt(mean((10^(spatial_pred.ridge)-10^(spatial_test$log10CHL4))^2))
rmse_spatial.ridge
R_squared.ridge = (sd(spar_pred))^2/(sd(spatial_train$log10CHL4))^2
R_squared.ridge
adj_R.ridge = 1-(1-R_squared.ridge)*((nrow(spatial_train)-1)/(nrow(spatial_train)-56))
adj_R.ridge
plot(10^(spatial_pred.ridge), 10^(spatial_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 8: Lasso regression with temporal data

```{r}
#LASSO MODEL
X = as.matrix(temporal_train[,-c(7:24)])
Y = as.matrix(temporal_train$log10CHL4)
lasso.mod = cv.glmnet(X, Y, alpha = 1, lambda = seq(0, 10, 0.01), nfolds = 10)
plot(lasso.mod)
plot(lasso.mod$glmnet.fit, "lambda", label = T)
lambda.lasso = lasso.mod$lambda.min
lasso.temporal = glmnet(X, Y, alpha = 1, lambda = lambda.lasso)
coef(lasso.temporal)
# Predict and compute MRSE
tem_pred = predict(lasso.temporal, as.matrix(temporal_train[,-c(7:24)]))
temporal_pred.lasso = predict(lasso.temporal, as.matrix(temporal_test[,-c(7:24)]))
rmse_temporal.lasso = sqrt(mean((10^(temporal_pred.lasso)-10^(temporal_test$log10CHL4))^2))
rmse_temporal.lasso
R_squared.lasso = (sd(tem_pred))^2/(sd(temporal_train$log10CHL4))^2
R_squared.lasso
adj_R.lasso = 1-(1-R_squared.lasso)*((nrow(temporal_train)-1)/(nrow(temporal_train)-32))
adj_R.lasso
plot(10^(temporal_pred.lasso), 10^(temporal_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 9: Ridge regression with temporal data

```{r}
library(MASS)
ridge.mod = lm.ridge(log10CHL4~., temporal_train[,-c(7:9, 11:24)], lambda = seq(0,10,0.01))
plot(ridge.mod)
best.lambda = as.numeric(names(which.min(ridge.mod$GCV)))
ridge.temporal = lm.ridge(log10CHL4~., temporal_train[,-c(7:9, 11:24)], lambda = best.lambda)
coef(ridge.temporal)
#Predict and compute RMSE
tem_pred = as.matrix(cbind(rep(1,nrow(temporal_train)),temporal_train[,-c(7:24)]))%*%as.vector(coef(ridge.temporal))
temporal_pred.ridge = as.matrix(cbind(rep(1,nrow(temporal_test)),temporal_test[,-c(7:24)]))%*%as.vector(coef(ridge.temporal))
rmse_temporal.ridge = sqrt(mean((10^(temporal_pred.ridge)-10^(temporal_test$log10CHL4))^2))
rmse_temporal.ridge
R_squared.ridge = (sd(tem_pred))^2/(sd(temporal_train$log10CHL4))^2
R_squared.ridge
adj_R.ridge = 1-(1-R_squared.ridge)*((nrow(temporal_train)-1)/(nrow(temporal_train)-32))
adj_R.ridge
plot(10^(temporal_pred.ridge), 10^(temporal_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 10: KNN regression with BATS data

```{r}
#KNN regression
library(caret)
knn.bats = knnreg(bats_train[,-c(7:24)], bats_train$log10CHL4)
bats_pred = predict(knn.bats, bats_train[,-c(7:24)])
bats_pred.knn = predict(knn.bats, bats_test[,-c(7:24)])
rmse_bats.knn = sqrt(mean((10^(bats_pred.knn)-10^(bats_test$log10CHL4))^2))
R_squared.knn = (sd(bats_pred))^2/(sd(bats_train$log10CHL4))^2
R_squared.knn
adj_R.knn = 1-(1-R_squared.knn)*((nrow(bats_train)-1)/(nrow(bats_train)-8))
adj_R.knn
cat("RMSE: ", rmse_bats.knn, "R-squared: ", R_squared.knn)
x = 1:nrow(bats_test)
plot(x,10^(bats_test$log10CHL4), col = 'red', type = 'l', lwd = 2, main = 'CHL 4 prediction', xlab = "Index", ylab = "Predicted CHL 4")
lines(x, 10^(bats_pred.knn), col ='blue', lwd =2)
legend("topright", legend = c("original value", "predicted value"), fill = c('red', 'blue'))
```

## Model 11: KNN regression with spatial data

```{r}
#KNN regression
library(caret)
knn.spatial = knnreg(spatial_train[,-c(57:74)], spatial_train$log10CHL4)
spatial_pred = predict(knn.spatial, spatial_train[,-c(57:74)])
spatial_pred.knn = predict(knn.spatial, spatial_test[,-c(57:74)])
rmse_spatial.knn = sqrt(mean((10^(spatial_pred.knn)-10^(spatial_test$log10CHL4))^2))
R_squared.knn = (sd(spatial_pred))^2/(sd(spatial_train$log10CHL4))^2
R_squared.knn
adj_R.knn = 1-(1-R_squared.knn)*((nrow(spatial_train)-1)/(nrow(spatial_train)-56))
adj_R.knn
cat("RMSE: ", rmse_spatial.knn, "R-squared: ", R_squared.knn)
x = 1:nrow(spatial_test)
plot(x,10^(spatial_test$log10CHL4), col = 'red', type = 'l', lwd = 2, main = 'CHL 4 prediction', xlab = "Index", ylab = "Predicted CHL 4")
lines(x, 10^(spatial_pred.knn), col ='blue', lwd =2)
legend("topright", legend = c("original value", "predicted value"), fill = c('red', 'blue'))
```

## Model 12: KNN regression with temporal data

```{r}
#KNN regression
library(caret)
knn.temporal = knnreg(temporal_train[,-c(7:24)], temporal_train$log10CHL4)
temporal_pred = predict(knn.temporal, temporal_train[,-c(7:24)])
temporal_pred.knn = predict(knn.temporal, temporal_test[,-c(7:24)])
rmse_temporal.knn = sqrt(mean((10^(temporal_pred.knn)-10^(temporal_test$log10CHL4))^2))
R_squared.knn = (sd(temporal_pred))^2/(sd(temporal_train$log10CHL4))^2
R_squared.knn
adj_R.knn = 1-(1-R_squared.knn)*((nrow(temporal_train)-1)/(nrow(temporal_train)-32))
adj_R.knn
cat("RMSE: ", rmse_temporal.knn, "R-squared: ", R_squared.knn)
x = 1:nrow(temporal_test)
plot(x,10^(temporal_test$log10CHL4), col = 'red', type = 'l', lwd = 2, main = 'CHL 4 prediction', xlab = "Index", ylab = "Predicted CHL 4")
lines(x, 10^(temporal_pred.knn), col ='blue', lwd =2)
legend("topright", legend = c("original value", "predicted value"), fill = c('red', 'blue'))
```

## Model 13: Elastic Net regression with BATS data

```{r}
library(glmnet)
library(caret)
bats_train_CHL4 = bats_train[,-c(7:9,11:24)]
elastic_net_bats = train(log10CHL4~., data = bats_train_CHL4, method = "glmnet")
elastic_net_bats
bats_pred.net = predict(elastic_net_bats, as.matrix(bats_test[,-c(7:24)]))
rmse_bats.net = sqrt(mean((10^(bats_pred.net)-10^(bats_test$log10CHL4))^2))
rmse_bats.net
bats_pred_train = predict(elastic_net_bats, as.matrix(bats_train[,-c(7:24)]))
R_squared.net = (sd(bats_pred_train))^2/(sd(bats_train$log10CHL4))^2
R_squared.net
adj_R.net = 1-(1-R_squared.net)*((nrow(bats_train)-1)/(nrow(bats_train)-8))
adj_R.net
plot(10^(bats_pred.net), 10^(bats_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 14: Elastic Net regression with spatial data

```{r}
# Elastic Net regression
library(glmnet)
library(caret)
spatial_train_CHL4 = spatial_train[,-c(57:59,61:74)]
elastic_net_spatial = train(log10CHL4~., data = spatial_train_CHL4, method = "glmnet")
elastic_net_spatial

spatial_pred.net = predict(elastic_net_spatial, as.matrix(spatial_test[,-c(57:74)]))
rmse_spatial.net = sqrt(mean((10^(spatial_pred.net)-10^(spatial_test$log10CHL4))^2))
rmse_spatial.net
spatial_pred_train = predict(elastic_net_spatial, as.matrix(spatial_train[,-c(57:74)]))
R_squared.net = (sd(spatial_pred_train))^2/(sd(spatial_train$log10CHL4))^2
R_squared.net
adj_R.net = 1-(1-R_squared.net)*((nrow(spatial_train)-1)/(nrow(spatial_train)-56))
adj_R.net
plot(10^(spatial_pred.net), 10^(spatial_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Model 15: Elastic Net regression with temporal data

```{r}
library(glmnet)
library(caret)
temporal_train_CHL4 = temporal_train[,-c(7:9,11:24)]
elastic_net_temporal = train(log10CHL4~., data = temporal_train_CHL4, method = "glmnet")
elastic_net_temporal
temporal_pred.net = predict(elastic_net_temporal, as.matrix(temporal_test[,-c(7:24)]))
rmse_temporal.net = sqrt(mean((10^(temporal_pred.net)-10^(temporal_test$log10CHL4))^2))
rmse_temporal.net
temporal_pred_train = predict(elastic_net_temporal, as.matrix(temporal_train[,-c(7:24)]))
R_squared.net = (sd(temporal_pred_train))^2/(sd(temporal_train$log10CHL4))^2
R_squared.net
adj_R.net = 1-(1-R_squared.net)*((nrow(temporal_train)-1)/(nrow(temporal_train)-32))
adj_R.net
plot(10^(temporal_pred.net), 10^(temporal_test$log10CHL4), xlab = "Predicted CHL4", ylab = "CHL 4")
abline(a = 0, b = 1, col = "red")
```

## Use model 2, 8, 13, 15 to predict on the validation set to assess the stability of each model

```{r}
# Model 2: Lasso regression on BATS data
bats_pred.lasso.valid = predict(lasso.bats, as.matrix(bats_valid[,-c(7:24)]))
rmse_bats.lasso.valid = sqrt(mean((10^(bats_pred.lasso.valid)-10^(bats_valid$log10CHL4))^2))
rmse_bats.lasso.valid

# Model 8: Lasso regression on temporal data
temporal_pred.lasso.valid = predict(lasso.temporal, as.matrix(temporal_valid[,-c(7:24)]))
rmse_temporal.lasso.valid = sqrt(mean((10^(temporal_pred.lasso.valid)-10^(temporal_valid$log10CHL4))^2))
rmse_temporal.lasso.valid

# Elastic Net regression on BATS data
bats_pred.net.valid = predict(elastic_net_bats, as.matrix(bats_valid[,-c(7:24)]))
rmse_bats.net.valid = sqrt(mean((10^(bats_pred.net.valid)-10^(bats_valid$log10CHL4))^2))
rmse_bats.net.valid

# Elastic Net regression on temporal data
temporal_pred.net.valid = predict(elastic_net_temporal, as.matrix(temporal_valid[,-c(7:24)]))
rmse_temporal.net.valid = sqrt(mean((10^(temporal_pred.net.valid)-10^(temporal_valid$log10CHL4))^2))
rmse_temporal.net.valid
```

## Use model 8 and 15 to predict all the CHL in all depths

### use Model 8 to predict all depths and draw plot for distribution of CHL in year 1994

```{r}
table_rmse = data.frame(matrix(data = NA, nrow = 17, ncol = 2))
table_prediction_8 = data.frame(matrix(data = NA, nrow = 1173, ncol = 1))
colnames(table_rmse) <- c("No.Depths", "RMSE of Lasso on temporal data")
table_rmse$No.Depths <- seq(2,18,1)
colnames(table_prediction_8) <- c("No.")

# Lasso with temporal data
for (i in 8:24) {
  X = as.matrix(temporal_data[,-c(7:24)])
  Y = as.matrix(temporal_data[,i])
  lasso.mod = cv.glmnet(X, Y, alpha = 1, lambda = seq(0, 10, 0.01), nfolds = 10)
  lambda.lasso = lasso.mod$lambda.min
  lasso.temporal = glmnet(X, Y, alpha = 1, lambda = lambda.lasso)
  # Predict and compute MRSE
  temporal_pred.lasso = predict(lasso.temporal, as.matrix(temporal_data[,-c(7:24)]))
  CHL_pred.lasso = 10^(temporal_pred.lasso)
  table_prediction_8 = cbind(table_prediction_8, CHL_pred.lasso)
  j = i-6
  colnames(table_prediction_8)[j] = paste("predict_depths", j) 
  rmse_temporal.lasso = sqrt(mean((10^(temporal_pred.lasso)-10^(temporal_data[,i]))^2))
  table_rmse[i-7,2] = rmse_temporal.lasso
}
table_prediction_8 = table_prediction_8[,-c(1)]
table_prediction_8 = cbind(table_prediction_8, temporal_data$year)
table_rmse # view table with RMSE of all depths
```

```{r}
# Draw CHL distribtion plot 
library(lattice)
library(raster)
library(rasterVis)
obs = as.matrix(c(1:69))
dp = as.matrix(c(1:17))
table_prediction_8.year = table_prediction_8[table_prediction_8$`temporal_data$year` == 1994,]
table_prediction_8.year = table_prediction_8.year[,c(17:1)]
col.l = colorRampPalette(c('red', 'orange', 'yellow', 'green', 'cyan', 'blue'))
#col.l = colorRampPalette(c('blue','cyan','yellow','orange','red'))
levelplot(as.matrix(table_prediction_8.year[,-c(18)]),row.values = obs, column.values = dp, col.regions = col.l, xlim = c(1:69), xlab = "Observation predicted", ylab = "Depths", main = "Distribution of CHL")
```

### Model 15: Elastic Net regression with temporal data

```{r}
table_rmse = data.frame(matrix(data = NA, nrow = 17, ncol = 2))
table_prediction_15 = data.frame(matrix(data = NA, nrow = 1173, ncol = 1))
colnames(table_rmse) <- c("No.Depths", "RMSE of Elastic Net with temporal data")
table_rmse$No.Depths <- seq(2,18,1)
colnames(table_prediction_15) <- c("No.")

# Elastic Net with temporal data
for (i in 8:24) {
  X = as.matrix(temporal_data[,-c(7:24)])
  Y = temporal_data[,i]
  elastic_net_temporal = train(X, Y, method = "glmnet")
  temporal_pred.net = predict(elastic_net_temporal, as.matrix(temporal_data[,-c(7:24)]))
  CHL_pred.net = 10^(temporal_pred.net)
  table_prediction_15 = cbind(table_prediction_15, CHL_pred.net)
  j = i-6
  colnames(table_prediction_15)[j] = paste("predict_depths", j) 
  rmse_temporal.net = sqrt(mean((10^(temporal_pred.net)-10^(temporal_data[,i]))^2))
  table_rmse[i-7,2] = rmse_temporal.net
}
table_prediction_15 = table_prediction_15[,-c(1)]
table_prediction_15 = cbind(table_prediction_15, temporal_data$year)
table_rmse # view table with RMSE of all depths
```

```{r}
# Draw CHL distribtion plot
library(lattice)
library(raster)
library(rasterVis)
obs = as.matrix(c(1:69))
dp = as.matrix(c(1:17))
table_prediction_15.year = table_prediction_15[table_prediction_15$`temporal_data$year` == 1994,]
table_prediction_15.year = table_prediction_15.year[,c(17:1)]
col.l = colorRampPalette(c('red', 'orange', 'yellow', 'green', 'cyan', 'blue'))
#col.l = colorRampPalette(c('blue','cyan','yellow','orange','red'))
levelplot(as.matrix(table_prediction_15.year[,-c(18)]),row.values = obs, column.values = dp, col.regions = col.l, xlim = c(1:69), xlab = "Observation predicted", ylab = "Depths", main = "Distribution of CHL")
```
